Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-01-27 komlos.tex
some related works

\begin{fact}
Let $f$ be a function with a corresponding discrepancy $K_m(f) = O(c/m)$. Let $F(q) = \sum_{i=1}^{n}f(x_i, q)$. There exists a coreset for $F$ of size $O(c/\eps)$ whose error is at most $\eps n$.
\end{fact}
\begin{fact}
Let $f$ be a function with a corresponding discrepancy $K_m(f) = O(c/\sqrt{m})$. Let $F(q) = \sum_{i=1}^{n}f(x_i, q)$. There exists a coreset for $F$ of size $O(c^2/\eps^2)$ whose error is at most $\eps n$.
\end{fact}

@@

\section{Related Work}
\begin{itemize}
\item coresets - motivation, some papers, streaming algorithms for coresets

\item Rademacher complexity \cite{Bartlett:2003:RGC:944919.944944} is a known method to estimate the generalization ability of an empirical risk minimizer (ERM) of a family of function over a finite set of i.i.d sampled points. A carefully selected set of points can potentially obtain a better generalization bound, and that is the usefulness of coresets in the context of machine learning. We are not aware of any analog to Rademacher complexity that aims to measure the generalization ability of a coreset of a fixed finite size. There are papers such as \cite{tolochinsky2018coresets} and references within that tie the coreset to the VC dimension of the family function, and the average sensitivity of the dataset. However, these relationships come up more as tools for constructing a coreset rather than a single complexity measure aimed to characterize some generalization ability.
%
% There is literature about the complexity of active learning for the supervised setting but that is out of scope as we deal with the setting where all labels are available, or the problem is unsupervised. For the complexity of coresets, the total sensitivity of the dataset and the VC dimension of the function class is often used as tools for construction \zk{cite something?}. However, to the best of our knowledge there is no clear characterization of the size of coresets based on some complexity measure of the dataset and function class.

@@

\item Kernel density estimation is a popular tool in data analysis aimed to estimate a continuous distribution with a finite set of points. Among other applications, this tool is used for outlier detection \cite{schubert2014generalized}, regression \cite{fan2018local}, and clustering \cite{rinaldo2010generalized}. A thorough survey could be found in \cite{silverman2018density}. Given a set of $n$ data points $x_1,\ldots,x_n$ and a query $q$, the density estimate of $q$ is the average $\frac{1}{n} \sum_i K(x_i,q)$ of the kernel evaluations with the data points. The task of finding a coreset for kernel density estimates consists of obtaining a set of points $z_1,\ldots,z_m$ with weights $w_1,\ldots,w_m$ for some $m \ll n$ such that for any query $q$ it holds that 
$\frac{1}{n} \sum_i K(x_i,q) = \frac{1}{m} \sum_i w_i K(z_i,q) \pm \eps$. For this task, the state-of-the-art is given by \cite{DBLP:journals/corr/abs-1802-01751}, achieving $m=\sqrt{d\log(1/\eps)}/\eps$ where $d$ is the dimension of the original data points. Their result holds for Lipchitz bounded positive semi-definite kernels. The result is constructive though it is polynomial rather than (quasi-)linear in the data size. The authors give an almost matching lower bound of $\sqrt{d}/\eps$ and pose an open question for closing the gap between the bounds.

>>>>

\begin{fact} \label{fct:eps1}
Let $f$ be a function with a corresponding discrepancy $K_m(f) = O(c/m)$. Let $F(q) = \sum_{i=1}^{n}f(x_i, q)$. There exists a coreset for $F$ of size $O(c/\eps)$ whose error is at most $\eps n$.
\end{fact}
\begin{fact}  \label{fct:eps2}
Let $f$ be a function with a corresponding discrepancy $K_m(f) = O(c/\sqrt{m})$. Let $F(q) = \sum_{i=1}^{n}f(x_i, q)$. There exists a coreset for $F$ of size $O(c^2/\eps^2)$ whose error is at most $\eps n$.
\end{fact}

@@

\section{Related Work}
\begin{itemize}
\item coresets - motivation, some papers, streaming algorithms for coresets.

\item Streaming coresets - MRL, etc. It is in fact known that for certain family types, a random sample of size $\log(1/\delta)/\eps^2$ is an $\eps$-coreset with probability $1-\delta$. For example, for families of VC dimension $v$, a sample of $(1/\eps^2)(v+\log(1/\delta))$ suffices \cite{talagrand1994sharper}.

\item Rademacher complexity \cite{Bartlett:2003:RGC:944919.944944} is a known method to estimate the generalization ability of an empirical risk minimizer (ERM) of a family of function over a finite set of i.i.d sampled points. A carefully selected set of points can potentially obtain a better generalization bound, and that is the usefulness of coresets in the context of machine learning. We are not aware of a complete analog to Rademacher complexity that aims to measure the generalization ability of a coreset of a fixed finite size. There are papers such as \cite{langberg2010universal, tolochinsky2018coresets} and references within that tie the coreset to the VC dimension of the family function, and the average sensitivity of the dataset. However, these relationships come up more as tools for constructing a coreset rather than a single complexity measure aimed to characterize some generalization ability. The connection between coresets and discrepancy have been known in the non-streaming setting. For example, \cite{phillips2009small} (Theorem 1.1) provide an analog to our Facts~\ref{fct:eps1},~\ref{fct:eps2}.
%
% There is literature about the complexity of active learning for the supervised setting but that is out of scope as we deal with the setting where all labels are available, or the problem is unsupervised. For the complexity of coresets, the total sensitivity of the dataset and the VC dimension of the function class is often used as tools for construction \zk{cite something?}. However, to the best of our knowledge there is no clear characterization of the size of coresets based on some complexity measure of the dataset and function class.

@@

\item Kernel density estimation is a popular tool in data analysis aimed to estimate a continuous distribution with a finite set of points. Among other applications, this tool is used for outlier detection \cite{schubert2014generalized}, regression \cite{fan2018local}, and clustering \cite{rinaldo2010generalized}. A thorough survey could be found in \cite{silverman2018density}. Given a set of $n$ data points $x_1,\ldots,x_n$ and a query $q$, the density estimate of $q$ is the average $\frac{1}{n} \sum_i K(x_i,q)$ of the kernel evaluations with the data points. The task of finding a coreset for kernel density estimates consists of obtaining a set of points $z_1,\ldots,z_m$ with weights $w_1,\ldots,w_m$ for some $m \ll n$ such that for any query $q$ it holds that 
$\frac{1}{n} \sum_i K(x_i,q) = \frac{1}{m} \sum_i w_i K(z_i,q) \pm \eps$. For this task, the state-of-the-art is given by \cite{DBLP:journals/corr/abs-1802-01751}, achieving $m=\sqrt{d\log(1/\eps)}/\eps$ where $d$ is the dimension of the original data points. Their result holds for Lipchitz bounded positive semi-definite kernels. The result is constructive though it is polynomial rather than (quasi-)linear in the data size. The authors give an almost matching lower bound of $\sqrt{d}/\eps$ and pose an open question for closing the gap between the bounds. The authors provide a simpler algorithm based on the Frank-Wolf method that achieves a $1/\eps^2$ sized coreset. It is in fact known (see \cite{lopez2015towards}, Theorem 1) that a sample of $\log(1/\delta)/\eps^2$ points gives a coreset w.p. $1-\delta$ for some kernel types.
